---
title: "Assignment 13"
author: "shawn schneidereit"
date: "10.2.2021"
output: html_document
---

**note** that the .Rmd can be found on my [personal repo](https://github.com/schneidereits/Quantitative_Methods_HU/tree/main/assignments) for the course

```{r, include= FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(viridis)
library(ggeffects)
library(FactoMineR)
library(MASS)
library(raster)
library(rasterVis)
library(GGally)
```


# A simple species distribution model.

The goal of this exercise is to build a simple statistical species distribution model for *Abies alba*. We will use a  Europe-wide tree species occurrence data set (Mauri et al., 2017) as reference and the Bioclim dataset as predictors.

Mauri, A., Strona, G., & San-Miguel-Ayanz, J. (2017). EU-Forest, a high-resolution tree occurrence dataset for Europe. Sci Data, 4, 160123


## 1) Calculate the count of *Abies alba* observations within each Bioclim grid cell (2 Points): 

We start by importing the dataset ("abiesalba.csv") and visualizing the observed occurrence of *Abies alba*. To make  this occurrence dataset spatially comparable to the predicted map that we will build later, I want you to create a raster of the occurrence counts at the spatial resolution of the bioclim dataset (bioclim_europe.tif). Check out the raster::rasterize() function to achieve this.

You can download the bioclim_europe.tif from: https://1drv.ms/u/s!AqNb6reoHXeZoq9JzbzkiwU_Gqu06w?e=wejkdx. The bioclim_europe raster has an equal-area projection (EPSG:3035, see: https://spatialreference.org/ref/epsg/etrs89-etrs-laea/). 
needed:
+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs 


```{r, warning=FALSE, message=F}
# Managing and manipulating vector data
library(sp)
# Reading and writing spatial data
library(rgdal)
# Reading and manipulating raster data
library(raster)
# Map visualization
library(mapview)
# Spatial analysis and modeling utilities
library(spatialEco)
# Managing and manipulating vector data
library(sf)

# import abies lat long csv
abies <- read_csv("~/Documents/humbolt/quantitative_methods/pc_lab/data/abiesalba.csv")

#convert to a simple features object
abies_sp <- SpatialPoints(abies, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# vis abies occurance
mapview(abies_sp)

# import eu bioclim data
clim <- stack("~/Documents/humbolt/quantitative_methods/pc_lab/data/bioclim_europe.tif")

# add variable names
names(clim) <- c("Annual Mean Temperature", "Mean Diurnal Range", "Isothermality", "Temperature Seasonality",
                 "Max Temperature of Warmest Month", "Min Temperature of Coldest Month", "Temperature Annual Range", "Mean Temperature of Wettest Quarter", "Mean Temperature of Driest Quarter", "Mean Temperature of Warmest Quarter", "Mean Temperature of Coldest Quarter", "Annual Precipitation", "Precipitation of Wettest Month", "Precipitation of Driest Month", "Precipitation Seasonality ", "Precipitation of Wettest Quarter", "Precipitation of Driest Quarter", "Precipitation of Warmest Quarter", "Precipitation of Coldest Quarter")

soil <- stack("~/Documents/humbolt/quantitative_methods/pc_lab/data/soil_europe_laea_10x10km.tif")

# creating a projection for eu bioclimatic data
clim_projection <- projection(clim)
clim_projection 

# transforming the abies data to match bio clim projection 
abies_sp <- spTransform(abies_sp, crs(clim_projection))

# generate a grid from the points extent
ext <- extent(abies_sp)
ext

#adding the restent to a new raster object
ras <- raster(ext)
ras 

# set resolution to 5000x5000, while keeping the original projection
res(ras) <- c(11200, 18500)
ras
# adding to projection (note this is the bioclim projection, not abies anymore)
projection(ras) <- projection(abies_sp)
ras # 5000x5000 resolution with the laea projection of the bioclim data

# rasterize the abie points and count the number of points in each cell
abies_ras <- raster::rasterize(abies_sp, clim, fun = "count", background = 0)
abies_ras
mapview(abies_ras)




```

## 2) Prepare a dataset combining response and predictor variables (3 Points): 

Before we can train our statistical model, we need to create a random sample from our response and predictor variables. The result should be a data.frame. Note, stack the bioclim and count raster before sampling. 

When sampling, you have to options: 1) a random sample of 20% of the observed grid-cells, OR 2) create a stratified sample containing 20% of the observations in each count class.

Option 2 gets extra points. Hint: You can combine dplyr::group_by() and dplyr::sample_frac()!

```{r}
# first we need harminize resolution and crop the climate data to have the same extent
res(clim) <- c(5000, 5000)
clim <- crop(clim, abies_ras)
plot(clim, y = 1)

# then we can stack 
abies_stack <- raster::stack(clim, abies_ras)

# option 1
# not best practice but hard coded to sample 20% of total cells (ncell)
abies_sample <- sampleRandom(abies_stack, size = 23219*.2)

# option 2
abies_sample <- abies_stack %>% 
  dplyr::group_by()

abies.df <- as.data.frame(abies_sample)
head(abies.df)

# rename variables 
names(abies.df) <- c("AnnualMeanTemperature", "MeanDiurnalRange", "Isothermality", "TemperatureSeasonality",
                 "MaxTemperatureofWarmestMonth", "MinTemperatureofColdestMonth", "TemperatureAnnualRange", "MeanTemperatureofWettestQuarter", "MeanTemperatureofDriestQuarter", "MeanTemperatureofWarmestQuarter", "MeanTemperatureofColdestQuarter", "AnnualPrecipitation", "PrecipitationofWettestMonth", "PrecipitationofDriestMonth", "PrecipitationSeasonality", "PrecipitationofWettestQuarter", "PrecipitationofDriestQuarter", "PrecipitationofWarmestQuarter", "PrecipitationofColdestQuarter", "Abies")

```

## 3) Build a model predicting the count of *Abies alba* (3 Points)

Use a GLM to predict the occurrence of *Abies alba* from bioclim data. You can hand-pick variables or use a variable selection technique. Your choice.

```{r}

m1 <- glm(data = abies.df, Abies ~ AnnualMeanTemperature + AnnualPrecipitation, family = "gaussian")
summary(m1)

# install.packages("MuMIn")
library(MuMIn)

# create semi-saturated model, with the broader variables based on my novice guess, I a priori think will be more relavent
m_saturated <- glm(data = abies.df, Abies ~ AnnualMeanTemperature + Isothermality + TemperatureSeasonality + TemperatureAnnualRange + AnnualPrecipitation + PrecipitationSeasonality, family = "gaussian")

summary(m_saturated)

options(na.action = "na.fail") # Required for dredge to run

model_dredge <- dredge(m_saturated, beta = F, evaluate = T, rank = AICc)

options(na.action = "na.omit") # set back to default

head(model_dredge)

nrow(model_dredge) # 64 models in total

top_model <- get.models(model_dredge, subset = 1)[[1]]
top_model

# Summarize top model
summary(top_model)

```

## 4) Create a map of the predicted *Abies alba* counts (2 Point)

```{r, warning=FALSE, message=F}

```



