---
title: "Assignment 14"
author: "Shawn Schneidereit"
date: "17.2.2021"
output: html_document
---

**note** that the .Rmd can be found on my [personal repo](https://github.com/schneidereits/Quantitative_Methods_HU/tree/main/assignments) for the course

```{r, include= FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(viridis)
library(ggeffects)
library(FactoMineR)
library(MASS)
library(raster)
library(rasterVis)
library(GGally)
```

## Climate station data for Germany

For this exercise we will use long-term average climate station data from the German Meteorological Service (DWD). Monthly and yearly average precipitation (mm) data are available for the period 1981 and 2010.

http://www.dwd.de/DE/leistungen/klimadatendeutschland/langj_mittelwerte.html

ftp://ftp-cdc.dwd.de/pub/CDC/

In this exercise, you are asked to create precipitation maps for the state of Brandenburg from weather station data using two different methods: 1) inverse distance weighted interpolation, and 2) ordinary kriging. Plot and compare the results/maps and semi-variograms. 

### 1. Create a SpatialPointsDataFrame.

Create a SpatialPointsDataFrame from the precipitation dataset `"nieder_1981-2010_aktStandort.csv"`. The precipitation dataset contains longitude (column `x`) and latitude (column `y`). The coordinate reference system is geographic WGS-84 (EPSG code 4326). In proj4 format this can be written as `"+init=epsg:4326"`. To analyze spatial distances you need to transform this dataset to an equal area projection, e.g. `"+init=epsg:3035"`. Note, use only the station data from Brandenburg (see `state` column).

```{r}

# Managing and manipulating vector data
library(sp)
# Reading and writing spatial data
library(rgdal)
# Reading and manipulating raster data
library(raster)
# Map visualization
library(mapview)
# Spatial analysis and modeling utilities
library(spatialEco)
# Managing and manipulating vector data
library(sf)
#
library(mapview)

#import data
rain <- read_csv("~/Documents/humbolt/quantitative_methods/pc_lab/data/nieder_1981-2010_aktStandort.csv")
head(rain)

# filter for brandenburg
rain <- rain %>% 
  filter(state == "Brandenburg") 

#creating seperate df
rain_df <- rain

coordinates(rain) <- ~x+y
proj4string(rain) <- CRS("+init=epsg:3035")


rain_sp <- SpatialPoints(rain, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))


# transforming the abies data to an equal area projection
rain_sp <- spTransform(rain_sp, crs("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))

mapview(rain_sp)
mapview(rain_sp, zcol="year")
```


### 2. Create a map of annual precipitation (column `year`) using Inverse Distance Weighted Interpolation.

```{r, fig.align='center', fig.height=4, fig.width=4}

xnew <- seq(rain@bbox[1, 1], rain@bbox[1, 2], length.out = 50)
ynew <- seq(rain@bbox[2, 1], rain@bbox[2, 2], length.out = 50)

gridnew <- expand.grid(xnew, ynew)

names(gridnew) <- c("x", "y")
coordinates(gridnew) <- ~x+y
proj4string(gridnew) <- CRS("+init=epsg:3035")
gridded(gridnew) <- TRUE


rain.grid <- SpatialPixelsDataFrame(points=gridnew[,c("x", "y")],
                                     data=gridnew[,(1:2)],
                                     proj4string = CRS(proj4string(rain_sp)))

mapview(rain.grid, zcol="year")

library(gstat)
zn.idw <- idw(year ~ 1, rain, rain_sp, idp = 2)
mapview(zn.idw, zcol='var1.pred', layer.name = "year")

```


### 3. Create a map of annual precipitation (column `year`) using ordinary kriging with constant intercept.

```{r}

```




