---
title: "Assignment 13"
author: "shawn schneidereit"
date: "10.2.2021"
output: html_document
---

**note** that the .Rmd can be found on my [personal repo](https://github.com/schneidereits/Quantitative_Methods_HU/tree/main/assignments) for the course

```{r, include= FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(viridis)
library(ggeffects)
library(FactoMineR)
library(MASS)
library(raster)
library(rasterVis)
library(GGally)
```


# A simple species distribution model.

The goal of this exercise is to build a simple statistical species distribution model for *Abies alba*. We will use a  Europe-wide tree species occurrence data set (Mauri et al., 2017) as reference and the Bioclim dataset as predictors.

Mauri, A., Strona, G., & San-Miguel-Ayanz, J. (2017). EU-Forest, a high-resolution tree occurrence dataset for Europe. Sci Data, 4, 160123


## 1) Calculate the count of *Abies alba* observations within each Bioclim grid cell (2 Points): 

We start by importing the dataset ("abiesalba.csv") and visualizing the observed occurrence of *Abies alba*. To make  this occurrence dataset spatially comparable to the predicted map that we will build later, I want you to create a raster of the occurrence counts at the spatial resolution of the bioclim dataset (bioclim_europe.tif). Check out the raster::rasterize() function to achieve this.

You can download the bioclim_europe.tif from: https://1drv.ms/u/s!AqNb6reoHXeZoq9JzbzkiwU_Gqu06w?e=wejkdx. The bioclim_europe raster has an equal-area projection (EPSG:3035, see: https://spatialreference.org/ref/epsg/etrs89-etrs-laea/). 
needed:
+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs 


```{r, warning=FALSE, message=F}
# Managing and manipulating vector data
library(sp)
# Reading and writing spatial data
library(rgdal)
# Reading and manipulating raster data
library(raster)
# Map visualization
library(mapview)
# Spatial analysis and modeling utilities
library(spatialEco)
# Managing and manipulating vector data
library(sf)
# import abies lat long csv
abies <- read_csv("~/Documents/humbolt/quantitative_methods/pc_lab/data/abiesalba.csv")
#convert to a simple features object
abies_sp <- SpatialPoints(abies, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
# vis abies occurance
mapview(abies_sp)
# import eu bioclim data
clim <- stack("~/Documents/humbolt/quantitative_methods/pc_lab/data/bioclim_europe.tif")
# add variable names
names(clim) <-  c("AnnualMeanTemperature", "MeanDiurnalRange", "Isothermality", "TemperatureSeasonality",
                 "MaxTemperatureofWarmestMonth", "MinTemperatureofColdestMonth", "TemperatureAnnualRange", "MeanTemperatureofWettestQuarter", "MeanTemperatureofDriestQuarter", "MeanTemperatureofWarmestQuarter", "MeanTemperatureofColdestQuarter", "AnnualPrecipitation", "PrecipitationofWettestMonth", "PrecipitationofDriestMonth", "PrecipitationSeasonality", "PrecipitationofWettestQuarter", "PrecipitationofDriestQuarter", "PrecipitationofWarmestQuarter", "PrecipitationofColdestQuarter")
soil <- stack("~/Documents/humbolt/quantitative_methods/pc_lab/data/soil_europe_laea_10x10km.tif")
# creating a projection for eu bioclimatic data
clim_projection <- projection(clim)
clim_projection 
# transforming the abies data to match bio clim projection 
abies_sp <- spTransform(abies_sp, crs(clim_projection))
# creating an object that containes the original abies extent
ext <- extent(abies_sp)
ext
##### obsolete as I transform abies_sp in the raster ransform step
# generate a grid from the points extent
#adding the restent to a new raster object
#ras <- raster(ext)
#ras 
# set resolution to 5000x5000, while keeping the original projection
#res(ras) <- c(11200, 18500)
#ras
# adding to projection (note this is the bioclim projection, not abies anymore)
#projection(ras) <- projection(abies_sp)
#ras # 5000x5000 resolution with the laea projection of the bioclim data
# rasterize the abie points and count the number of points in each cell
abies_ras <- raster::rasterize(abies_sp, clim, fun = "count", background = 0)
# crop to match orignal abies extent 
abies_ras <- crop(abies_ras, ext)
abies_ras
mapview(abies_ras)
```

## 2) Prepare a dataset combining response and predictor variables (3 Points): 

Before we can train our statistical model, we need to create a random sample from our response and predictor variables. The result should be a data.frame. Note, stack the bioclim and count raster before sampling. 

When sampling, you have to options: 1) a random sample of 20% of the observed grid-cells, OR 2) create a stratified sample containing 20% of the observations in each count class.

Option 2 gets extra points. Hint: You can combine dplyr::group_by() and dplyr::sample_frac()!

```{r}
# first we need harminize resolution and crop the climate data to have the same extent
clim <- crop(clim, abies_ras)
plot(clim, y = 1)
# then we can stack 
abies_stack <- raster::stack(clim, abies_ras)
# option 1
# not best practice but hard coded to sample 20% of total cells (ncell)
abies_sample <- sampleRandom(abies_stack, size = 23219*.2)
abies.df <- as.data.frame(abies_sample) 
head(abies.df)
# rename variables 
names(abies.df) <- c("AnnualMeanTemperature", "MeanDiurnalRange", "Isothermality", "TemperatureSeasonality",
                 "MaxTemperatureofWarmestMonth", "MinTemperatureofColdestMonth", "TemperatureAnnualRange", "MeanTemperatureofWettestQuarter", "MeanTemperatureofDriestQuarter", "MeanTemperatureofWarmestQuarter", "MeanTemperatureofColdestQuarter", "AnnualPrecipitation", "PrecipitationofWettestMonth", "PrecipitationofDriestMonth", "PrecipitationSeasonality", "PrecipitationofWettestQuarter", "PrecipitationofDriestQuarter", "PrecipitationofWarmestQuarter", "PrecipitationofColdestQuarter", "Abies")
# option 2
abies.df <- as.data.frame(abies_stack)
abies_sample <- abies.df %>% 
  dplyr::group_by(layer) %>% 
  dplyr::sample_frac(layer, 20) %>% 
  mutate(Abies = layer)
```

## 3) Build a model predicting the count of *Abies alba* (3 Points)

Use a GLM to predict the occurrence of *Abies alba* from bioclim data. You can hand-pick variables or use a variable selection technique. Your choice.

```{r}
ggplot(abies_sample, aes(x = Abies)) +
         geom_density()
m1 <- glm(data = abies_sample, Abies ~ AnnualMeanTemperature + AnnualPrecipitation, family = "poisson")
summary(m1)
# install.packages("MuMIn")
library(MuMIn)
# create semi-saturated model, with the broader variables based on my novice guess, I a priori think will be more relavent
m_saturated <- glm(data = abies_sample, Abies ~ AnnualMeanTemperature + Isothermality + TemperatureSeasonality + TemperatureAnnualRange + AnnualPrecipitation + PrecipitationSeasonality, family = "poisson")
summary(m_saturated)
options(na.action = "na.fail") # Required for dredge to run
model_dredge <- dredge(m_saturated, beta = F, evaluate = T, rank = AICc)
options(na.action = "na.omit") # set back to default
head(model_dredge)
nrow(model_dredge) # 64 models in total
top_model <- get.models(model_dredge, subset = 1)[[1]]
top_model
# Summarize top model
summary(top_model)
# psudo R^2
1- summary(top_model)$deviance / summary(top_model)$null.deviance
```

## 4) Create a map of the predicted *Abies alba* counts (2 Point)


```{r, warning=FALSE, message=F}
library(gridExtra)
#predict to a raster
abies_predited <- predict(abies_stack, top_model, progress='text')
# comparing orginal and predicted map
mapview(abies_ras)
mapview(abies_predited)
```