---
title: "Assignment Lab Session 2"
author: "Shawn Schneidereit"
date: "11.11.2020"
output: html_document
---

### Steps to complete your assignment:

1. Complete the tasks. Write your R code in the designated areas (you can send the code to the console to test your code).
2. Create an HTML file in R-Studio: | Menu | File | Knit Document (alternatively: little Knit button in script header)
3. RENAME your html file to "Assignment_Lab02_YourLastName.html" (replace your last name)
4. Upload your assignment (.html file) to Moodle


### 1. Matrices (2 Points)

Convert vector x into a matrix m with 5 rows und 8 columns.

```{r}
x <- c(2,2,2,2,2,2,2,2,2,2,2,8,8,8,8,1,2,3,4,5,6,7,8,9,10,6,5,4,3,2,1,56,56,56,56,8,8,8,8,8)

m <- matrix(x, nrow = 5, ncol = 8)

m
```

Extract the third column from matrix m.

```{r}
m[,3]

```

From matrix m, extract the element from the fourth row and second column.

```{r}
m[4,2]


```

From matrix m, extract the 7ths and 8ths column.

```{r}
m[,7:8]


```

Convert matrix m into a data.frame names 'mm' and extract the 3rd column

```{r}
mm <- data.frame(m)
mm[,3]
```

### 2. Data frames (2 Points)

Download the airquality data set from Moodle and import it into R.

```{r}
library(readr)
airquality <- read_csv("~/Documents/humbolt/quantitative_methods/pc_lab/data/airquality_output.csv")
```

Answer the following questions:

- What is the concentration of ozone on May 20th?
- What is the mean temperature on July 16th?
- How many days had an ozone concentration greater than 50 ppm?

```{r}
# What is the concentration of ozone on May 20th?
# may 20th = month 5, day 20; found in row 20
airquality[20, "Ozone"] # answer: 11


# What is the mean temperature on July 16th?
# july 16 = month 7, day 20; found in row 77
airquality[77, "Temp"] # answer: 81


# How many days had an ozone concentration greater than 50 ppm?
length(airquality["Ozone" > 50])  # answer: 81

```

### 3. Manipulate data (3 Points)

Convert column `Month` (numeric) to a factor using the months names as factor labels ("May", "June", ...).

```{r}
airquality <- read_csv("~/Documents/humbolt/quantitative_methods/pc_lab/data/airquality_output.csv")


airquality$Month[airquality$Month == 5] <- "May"
airquality$Month[airquality$Month == 6] <- "June"
airquality$Month[airquality$Month == 7] <- "July"
airquality$Month[airquality$Month == 8] <- "August"
airquality$Month[airquality$Month == 9] <- "September"
airquality$Month <- as.factor(airquality$Month)


```

Use the function `paste()` to create a new column called 'Date' of the following format: 'Month-Day', 'May-1', 'May-2', 'May-3', ...

```{r}
airquality$Date <- paste(airquality$Month, airquality$Day, sep = "-")

```

Create two new variables ('Temp.C' and 'Wind.ms') that show temperature in Celsius and wind speed in m/s (instead of Fahrenheit und mph). Show the first five records for each new variable.

```{r}
airquality$Temp.C <- round((airquality$Temp - 32)/1.8, digits = 2)

airquality$Wind.ms <- airquality$Wind / 2.237 # divide by 2.237 to convert to mph to m/s

head(airquality[,c("Temp.C", "Wind.ms")])

```

Write the data to a file called 'airquality2.csv' using csv format.

```{r}
write_csv(airquality, path = "~/Documents/humbolt/quantitative_methods/pc_lab/data/airquality2.csv")
```

### 4. Loading new data (3 points)

Download the dataset 'treering.csv' from Moodle. Inspect it in either Excel or a text editor (Notepad++ for Windows or TextWrangler for Mac). Try loading the dataset into a data frame.

The dataset contains the width of treerings measures at several plots.

```{r}
treering <- read_csv("~/Documents/humbolt/quantitative_methods/pc_lab/data/treering.csv")
```

Drop all rows containing at least one `NA` value.

```{r}
treering <- na.omit(treering)
```

What is the mean tree-ring width for plot X40B?

```{r}
mean(treering$X_40_B) # answer 1.221236

```

