---
title: "Asignment Week 1"
author: "Shawn Schneidereit"
date: "2.11.2020"
output: html_document
---

## **note** that the .Rmd can be found on my [personal repo](https://github.com/schneidereits/Quantitative_Methods_HU/blob/main/pc_lab/assignment_week_01.Rmd) for the corse


```{r, include= FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(viridis)
library(ggeffects)
library(GGally)
library(FactoMineR) # pca?
library(lme4)
```

# Constraining the effects of topography and climate on climate change sensitivity of glaciers in Tibet

In this analysis I attempted to justify all modelling structure and variable selection decisions made. As I quite unfamiliar with climatic modelling and glacier dynamics, some of these decisions may we sub-optimal or completely wrong. Yet, these decisions where made to the best of my abilites, without going over the top on background research, overall my models should have a cohernt stucture and suffienct work flow documentaion. 
 
Multiple sources confirm that overall there is a lack of knowledge on climatic data and their influence on the TP, due to a lack of Permanente meteorological weather stations in the TP, with the influence of Monsoon being aknoweleged (Maussion et. al., 2014).

### 1.Data import and wrangeling


```{r}
glacier <- read_csv("~/Documents/humbolt/quantitative_methods/assignments/glacier_sensitivity.csv") %>%
  na.exclude 


#ggplot(data=glacier, aes_string(x = id, y = "")) + geom_line()

colnames(glacier) 

(corr_plot <- ggpairs(data = glacier, columns = c(2:9,22:25,38:39,52:53,66:74)))
        




```

### Q1.	Which are the most important topographical and climatological drivers of glacier equilibrium line altitude (ELA)?

#### Mixed effects model with random effects

In this first section, I chose to create a Mixed effects model with random effects (lme4 package), where all variables where chosen based on a priori knowelge gained from the readings and the exploratory correlation plot. The methodoological advantage of this approach is that through including the random effects correlations between data coming from specific morphology types and geographical orientations can be accounted for, by treating them as a grouping factor. While this prevents us from explicitly assessng the impacts of morphology and geographic orientation, it still allows us to see how it is influencing the observed patterns in other varibales. In the final section of this analysis (the automated variable selection) morphology and geographic orientation are left as fixed effects, so that I have them once as a fixed and once as a random effect, to allow for compairson between the two secarios. 

```{r}

glacier_scaled <- glacier %>% 
  # adding a new variable for temperature variability; a factor mentioned to be important in multiple papers
  mutate(#T_variability = (mean(T_MAX_mean_monsoon, T_MAX_mean_not_monsoon)-(mean(T_MIN_mean_monsoon, T_MIN_mean_not_monsoon))),
          # scaling all varibales used in modelling
         length = scale(length, center = T, scale = T),
         area = scale(area, center = T, scale = T),
         P_snow = scale(P_snow, center = T, scale = T),
         P_year = scale(P_year, center = T, scale = T),
         P_monsoon = scale(P_monsoon, center = T, scale = T),
         P_not_monsoon = scale(P_not_monsoon, center = T, scale = T),
         T_MIN_mean_monsoon = scale(T_MIN_mean_monsoon, center = T, scale = T),
         T_MIN_mean_not_monsoon = scale(T_MIN_mean_not_monsoon, center = T, scale = T),
         T_MAX_mean_monsoon = scale(T_MAX_mean_monsoon, center = T, scale = T),
         T_MAX_mean_not_monsoon = scale(T_MAX_mean_not_monsoon, center = T, scale = T),
        # T_variability = scale(T_variability, center = T, scale = T),
         T_mean_mea.yr = scale(T_mean_mea.yr, center = T, scale = T),
         T_mean_monsoon = scale(T_mean_monsoon, center = T, scale = T),
         T_mean_not_monsoon = scale(T_mean_not_monsoon, center = T, scale = T),
         Slope_min = scale(Slope_min, center = T, scale = T),
         Slope_max = scale(Slope_max, center = T, scale = T),
         Slope_mean = scale(Slope_mean, center = T, scale = T),
         Elev_min = scale(Elev_min, center = T, scale = T),
         Elev_max = scale(Elev_max, center = T, scale = T),
         Elev_mean = scale(Elev_mean, center = T, scale = T))

m_lme4 <- lmer(ELA ~ debris_cov + length + area + P_snow + T_mean_mea.yr + Slope_mean + Elev_max + Elev_mean + (1|morph_type) + (1|orientation), data = glacier_scaled) 
summary(m_lme4)

library(sjPlot)
library

plot_model(m_lme4, type = "re", show.values = TRUE)
```


#### Q1b PCA 
```{r}
# pca
pca <- glacier[c(2:9,22:25,38:39,52:53,66:74)] %>% 
  dplyr::select(-orientation) # need to remove as it is a character
str(pca)

# plot of pc1 and pc2
res.pca <- PCA(pca[,c(1:24)], scale.unit = TRUE, ncp = 5, graph = TRUE)
# based on this PCA we can intially see that ELA is closly aligned with PC1 and has a strong positive orientation with elev mean, while being diamertically opposed to the agregated temperature and precipitation measures. Geo-physical conditions (aside from elev mean on PC1) tend to be ordinated on PC2 with the strongers factors being glacier area, length, and summit height/max elevation (the same redundent metric...). Less promising is the fact that dELA does not load on either PC1 or PC2, and only accounts for a small amount of variance, as it has a small eigenvector.


#plot of pc1 and pc3
res.pca <- PCA(pca[,c(1:24)], scale.unit = TRUE, ncp = 5, graph = TRUE, axes = c(1,3))
# based on this plot of PC1 and PC2, we can see that snow precipitation is the main variable contributing to PC3s explained variation. Furthermore the ordination of variables on PC1 are only slightly altered in compairoson to the plot of PC1 and PC2, with one notable shift being that dELA closely ordinates in the direction of PC1 (but as before only has a small eigenvector)  


pca <- prcomp(pca, scale. = T)
summary(pca)



var_exp <- data.frame(pc = c(2:6,8:9,22:25,38:39,52:53,66:74),
                      var_exp = pca$sdev^2 / sum(pca$sdev^2))

# add the variances cumulatively
var_exp$var_exp_cumsum <- cumsum(var_exp$var_exp)
var_exp

ggplot(var_exp) +
  geom_bar(aes(x = pc, y = var_exp), stat = "identity") +
  geom_line(aes(x = pc, y = var_exp_cumsum)) +
  theme_classic()


```

### Q2.	Can we explain ELA changes since the Little Ice Age with topographical and climatological variables?

```{r}

```

### 4. 

```{r}

```

### 5. 

```{r}

```

