---
title: "Asignment Week 1"
author: "Shawn Schneidereit"
date: "2.11.2020"
output: html_document
---

## **note** that the .Rmd can be found on my [personal repo](https://github.com/schneidereits/Quantitative_Methods_HU/blob/main/pc_lab/assignment_week_01.Rmd) for the corse


```{r, include= FALSE}
library(readr)
library(viridis)
library(ggeffects)
library(GGally)
library(FactoMineR) # pca?
library(lme4)
library(tidyverse)
library(dotwhisker)
```

# Constraining the effects of topography and climate on climate change sensitivity of glaciers in Tibet

In this analysis I attempted to justify all modelling structure and variable selection decisions made. As I quite unfamiliar with climatic modelling and glacier dynamics, some of these decisions may we sub-optimal or completely wrong. Yet, these decisions where made to the best of my abilites, without going over the top on background research, overall my models should have a cohernt stucture and suffienct work flow documentaion. 
 
Multiple sources confirm that overall there is a lack of knowledge on climatic data and their influence on the TP, due to a lack of Permanente meteorological weather stations in the TP, with the influence of Monsoon being aknoweleged (Maussion et. al., 2014).

### 1.Data import and wrangeling


```{r}
glacier <- read_csv("~/Documents/humbolt/quantitative_methods/assignments/glacier_sensitivity.csv") %>%
  na.exclude 

glacier_scaled <- glacier %>% 
  # adding a new variable for temperature variability; a factor mentioned to be important in multiple papers
  mutate(T_variability = ((T_MAX_mean_monsoon + T_MAX_mean_not_monsoon)-(T_MIN_mean_monsoon + T_MIN_mean_not_monsoon) /2),
          # scaling all varibales used in modelling
         length = scale(length, center = T, scale = T),
         area = scale(area, center = T, scale = T),
         P_snow = scale(P_snow, center = T, scale = T),
         P_year = scale(P_year, center = T, scale = T),
         P_monsoon = scale(P_monsoon, center = T, scale = T),
         P_not_monsoon = scale(P_not_monsoon, center = T, scale = T),
         T_MIN_mean_monsoon = scale(T_MIN_mean_monsoon, center = T, scale = T),
         T_MIN_mean_not_monsoon = scale(T_MIN_mean_not_monsoon, center = T, scale = T),
         T_MAX_mean_monsoon = scale(T_MAX_mean_monsoon, center = T, scale = T),
         T_MAX_mean_not_monsoon = scale(T_MAX_mean_not_monsoon, center = T, scale = T),
         T_variability = scale(T_variability, center = T, scale = T),
         T_mean_mea.yr = scale(T_mean_mea.yr, center = T, scale = T),
         T_mean_monsoon = scale(T_mean_monsoon, center = T, scale = T),
         T_mean_not_monsoon = scale(T_mean_not_monsoon, center = T, scale = T),
         Slope_min = scale(Slope_min, center = T, scale = T),
         Slope_max = scale(Slope_max, center = T, scale = T),
         Slope_mean = scale(Slope_mean, center = T, scale = T),
         Elev_min = scale(Elev_min, center = T, scale = T),
         Elev_max = scale(Elev_max, center = T, scale = T),
         Elev_mean = scale(Elev_mean, center = T, scale = T))




#ggplot(data=glacier, aes_string(x = id, y = "")) + geom_line()

colnames(glacier) 

(corr_plot <- ggpairs(data = glacier, columns = c(2:9,22:25,38:39,52:53,66:74)))
#corplot?
corrplot.mixed(cor1[,2:9,22:25,38:39,52:53,66:74], lower.col = “black”, number.cex = .7)

        




```

### Q1.	Which are the most important topographical and climatological drivers of glacier equilibrium line altitude (ELA)?

#### Mixed effects model with random effects

In this first section, I chose to create a Mixed effects model with random effects (lme4 package), where all variables where chosen based on a priori knowelge gained from the readings and the exploratory correlation plot. The methodoological advantage of this approach is that through including the random effects, correlations between data coming from specific morphology types and geographical orientations can be accounted for, by treating them as a grouping factor. While this prevents us from explicitly assessng the impacts of morphology and geographic orientation, it still allows us to see how it is influencing the observed patterns in other varibales. In the final section of this analysis (the automated variable selection) morphology and geographic orientation are left as fixed effects, so that I have them once as a fixed and once as a random effect, to allow for compairson between the two secarios. As the Distribution of ELA had a slight negative skew, I also tried running a GLM with a inverse gaussian distribution (link = "1/mu^2"), but encoured the error that the "PIRLS step-halvings failed to reduce deviance in pwrssUpdate", to which I did not find a solution. 

```{r}

hist(glacier_scaled$ELA) # there is a negative skew to the data, but overall a gaussian distibution looks appropriate

m_lme4 <- lmer(ELA ~ debris_cov + length + area + P_snow + P_monsoon + P_not_monsoon +  T_mean_mea.yr +  T_variability +                      Slope_mean + Elev_max + Elev_mean + (1|morph_type) + (1|orientation), 
               data = glacier_scaled) 
summary(m_lme4)

# R squared
r.squaredGLMM(m_lme4)
# The R²(C) is 90.78 %, where the C represents the Conditional R_GLMM² and can be interpreted as a variance explained by the entire model, including both fixed and random effects

# check out variable co-lineararity 

plot(m_lme4)
# the residual vs fitted plot shows a mostly flat and even distribution of points, with greater diviances seen at the lower spectrum of points, likely due to the negative skew described above

qqnorm(resid(m_lme4))
qqline(resid(m_lme4))
# When checking the QQ-plot it is seen that there is a left tail at lower bonds and a right tail at the upper bonds. Following an ad hoc transformation of the dependant variable ELA, this observed trend only showed a very marginal improvment, so I decided to keep ELA in its untransformed state.  




dwplot(m_lme4) +
   geom_vline(xintercept = 0, colour = "grey60", linetype = 2)
```

```{r}

hist(glacier_scaled$dELA) # there is a stronger skew to the data. While again an inverse Gaussian distribution seems approaprite, this was also not possible. Instead I compaired a the model with a normal Gaussian and a Possion distribution and found that the Possion had a better distribution of residuals.

m_lme4_d <- glmer(dELA ~ debris_cov + length + area + P_snow + P_monsoon + P_not_monsoon +  T_mean_mea.yr +  T_variability +                      Slope_mean + Elev_max + Elev_mean + (1|morph_type) + (1|orientation), 
                family = poisson  ,data = glacier_scaled) 
summary(m_lme4_d)

# R squared
r.squaredGLMM(m_lme4_d)
# The R²(C) is 98.65 %, where the C represents the Conditional R_GLMM² and can be interpreted as a variance explained by the entire model, including both fixed and random effects

# check out variable co-lineararity 

plot(m_lme4_d)
# the residual vs fitted plot shows a mostly flat and even distribution of points, with some outlying point at intermidiate values

qqnorm(resid(m_lme4_d))
qqline(resid(m_lme4_d))
# When checking the QQ-plot it is seen that there is slight tail at the upper bound, but overall this is a significant improvement over the gaussian distribution.


dwplot(m_lme4_d) +
   geom_vline(xintercept = 0, colour = "grey60", linetype = 2)
```


#### Q1b PCA 
```{r}
# pca # need to redo analysis/description of pca plots
pca <- glacier_scaled[c(4:9,22:25,38:39,52:53,66:75)] %>% 
  dplyr::select(-orientation) # need to remove as it is a character
str(pca)

# plot of pc1 and pc2
res.pca <- PCA(pca[,c(1:22)], scale.unit = TRUE, ncp = 5, graph = TRUE) 
# based on this PCA we can intially see that elev mean is closly aligned with PC1 and has a strong negative orientation, while being diamertically opposed to the agregated temperature and precipitation measures. Geo-physical conditions (aside from elev mean on PC1) tend to be ordinated on PC2 with the strongers factors being glacier area, length, and summit height/max elevation (the same redundent metric...). 


#plot of pc1 and pc3
res.pca <- PCA(pca[,c(1:22)], scale.unit = TRUE, ncp = 5, graph = TRUE, axes = c(1,3))
# based on this plot of PC1 and PC2, we can see that snow precipitation is the main variable contributing to PC3s explained variation. Furthermore the ordination of variables on PC1 are only slightly altered in compairoson to the plot of PC1 and PC2. 


pca <- prcomp(pca, scale. = T)
summary(pca)
pca$rotation[,1:6]
# based on this roation, I will make the following assumption for which variables are analogs to each PC
# PC1: All tempurature variables (both monsoon & non-monsoon), tempurature variability, and elevation mean
# PC2: Geophysical conditions such as min/max elevation, debris cover, catchment length and area, and max slope
# PC3: All precipitation variables 
# PC4: Slope, and to a far lesser degree catchment area, length and elevation
# PC5: Morphology type 


var_exp <- data.frame(pc = c(1:23),
                      var_exp = pca$sdev^2 / sum(pca$sdev^2))

# add the variances cumulatively
var_exp$var_exp_cumsum <- cumsum(var_exp$var_exp)
var_exp

ggplot(var_exp) +
  geom_bar(aes(x = pc, y = var_exp), stat = "identity") +
  geom_line(aes(x = pc, y = var_exp_cumsum)) +
  theme_classic()

# Following the Elbow method, it seems like 3-5 PCs seems like an approapriate cutoff point. Despite being more difficult to interpret the final results, I decided to include 5 PCs, due to PCs 4 and 5 increasing the overall variance explained from ~80% to 90%. This also helps to ensure that the large number of variables in the model are captured in the PCs, especially as morphology type almost exclusively loads on PC5.

# scores
pca_scores <- data.frame(pca$x)
head(pca_scores)

pca_scores_df <- cbind(glacier_scaled, pca_scores)
ggpairs(pca_scores_df[,c(4:6,8:9,22:25,38:39,52:53,66:75,76:80)]) # first 5 Pcs

m_pca <- glm(ELA ~ PC1 + PC2 + PC3 + PC4 + PC5, data = pca_scores_df)
summary(m_pca)
# Here we see that PCs 1,2, and 4 have the strongest effect size and large predictive power for ELA. Based on the caracterization of which variables are represented by each PC, this means that tempurature, preciptiation, and slope (and to a smaller extend catchment length, area, and elevation) are the most important variables. Out of the three PCs, PC1 (tempurature variables), are by far the most important with a effect size of -65.17 (Std. error 0.88). PCs 2 (precipitation) and 4 (slope) have effect sizes of 17.78 (Std. error 1.35) and -24.04 (Std. error 2.27) respectively. 

# R squared
r.squaredGLMM(m_pca)
# The R² is 87.12 %

plot(m_pca)
# the residual vs fitted plot shows a mostly flat and even distribution of points, with some outlying point at intermidiate values. The QQ-plot has skewed tails at both the lower and upper bounds and is substaintial for larger values


dwplot(m_pca) +
   geom_vline(xintercept = 0, colour = "grey60", linetype = 2)


# dELA model
m_pca_d <- glm(dELA ~ PC1 + PC2 + PC3 + PC4 + PC5, family = "poisson", data = pca_scores_df)
summary(m_pca)

# R squared
r.squaredGLMM(m_pca_d)
# The R² is 76.44 %

plot(m_pca_d)
# the residual vs fitted plot shows a mostly flat and even distribution of points, with some outlying point at intermidiate values


dwplot(m_pca_d) +
   geom_vline(xintercept = 0, colour = "grey60", linetype = 2)
```



### Automated variable selection 

For the automated variable variable I focused on variables that I previously found to be strong predictors of ELA, aswell as the deagrigated forms of temperature and precipitation metrics. While potential interaction between temperature and precipitation variables are likely to exist, these were not included in the saturated model, as just including variables with no interactions already brought my computer to the edges of its computational limits. 

```{r}

library(MuMIn)
# create semi-saturated model, with the broader variables based on my novice guess, I a priori think will be more relavent
m_saturated <-
  glm( ELA ~ debris_cov  + morph_type + orientation + length  + area  + P_snow + P_year + P_monsoon + P_not_monsoon +                    T_mean_mea.yr + T_mean_monsoon +   T_mean_not_monsoon + Slope_min + Slope_max + Slope_mean + Elev_min + Elev_max +              Elev_mean + T_variability, # + T_MIN_mean_monsoon + T_MIN_mean_not_monsoon + T_MAX_mean_monsoon  T_MAX_mean_not_monsoon ; taken out due to computational constraints
  data = glacier_scaled,
  family = "gaussian") 

summary(m_saturated)
options(na.action = "na.fail") # Required for dredge to run
model_dredge <- dredge(m_saturated, beta = F, evaluate = T, rank = AICc)
options(na.action = "na.omit") # set back to default
head(model_dredge)
nrow(model_dredge) # 64 models in total
top_model <- get.models(model_dredge, subset = 1)[[1]]
top_model
# Summarize top model
summary(top_model)
# psudo R^2
1- summary(top_model)$deviance / summary(top_model)$null.deviance

```

### Q2.	Can we explain ELA changes since the Little Ice Age with topographical and climatological variables?


```{r}

```

### 5. 

```{r}

```

