---
title: "Asignment Week 1"
author: "Shawn Schneidereit"
date: "2.11.2020"
output: html_document
---

## **note** that the .Rmd can be found on my [personal repo](https://github.com/schneidereits/Quantitative_Methods_HU/blob/main/pc_lab/assignment_week_01.Rmd) for the corse

### 1. Create a vector from a series of numbers:

2;2;2;2;2;2;2;2;2;2;2;8;8;8;8;1;2;3;4;5;6;7;8;9;10;6;5;4;3;2;1;56;56;56;56;8;8;8;8;8

From the above number series construct a vector `x` using the functions `c()`, `rep()` and `seq()`. Note, you MUST use all three functions! Read the slides from the lecture carefully to answer the questions.

```{r}
x <- c(rep(2, times = 11),rep(8, times = 4), seq(1, 10), seq(6, 1), rep(56, times = 4), rep(8, times = 5))
x


```

### 2. How many elements has vector `x`?

```{r}
x <- c(2,2,2,2,2,2,2,2,2,2,2,8,8,8,8,1,2,3,4,5,6,7,8,9,10,6,5,4,3,2,1,56,56,56,56,8,8,8,8,8)
length(x)

```

### 3. Extract the 12th element from vector `x`.

```{r}
x[12]
```

### 4. Extract the elements 20-27 from vector `x`.

```{r}
x[20:27]
```

### 5. Extract the 12th and 20-27th element from vector `x`.

```{r}
x[c(12, 20:27)]
```

### 6. Extract all but the 5th element from vector `x` and assign the new vector to a variable `y`. Print the content of `y`.

```{r}
y <- x[-5]
y
```

### 7. Write a logical expression that assesses for each element `i` in vector `x` whether `i` is equal to `56`. The result should be a logical vector of `TRUE`s and `FALSE`s that has the same number of elements as `x`.

```{r}
x == 56

# technically we did not cover this in the lecture, but this is a ifelse statement approach
ifelse(x == 56, TRUE, FALSE)
```

### 8. Use the logical expression to replace all values `56` in vector `x` with the value `52`. Print the content of `x`.

```{r}
x[x == 56] <- 52
x

# technically we did not cover this in the lecture, but this is a ifelse statement approach
ifelse(x == 56, 52, x)
x
```

### 9. Replace all elements in vector `x` that are less than `5` or greater than `50` with the value `NA`. Use a logical expression, and print the result.

```{r}

x[x < 5 | x > 50] <- NA
x

# technically we did not cover this in the lecture, but this is a ifelse statement approach
ifelse(x < 5 | x > 50, NA, x)
x
```

### 10. Add `5` to each element in vector `x`.

```{r}
x + 5
x
```
